---
title: "Defensive Behaviour of M.separata in Response to Different Parasitoid Species"
author: "Blessing_Amira_Aizada_Sami_Ali"
format: html
editor: visual
---

# Extracting the Data

```{r}
library(dplyr)
library(ggplot2)
library(readxl)
library(sandwich)
library(lmtest)
```

```{r}
# Set file path (replace with your actual file path)
file_path <- "data/parasitoid.xlsx"

d0 <- read_excel(file_path)
```

# A logistic regression analysis was conducted to evaluate the relationship between host weight and thrashing frequency as predictors of dropping behavior (Drop vs No Drop) in M. mediator. The model estimated the likelihood of dropping, and predicted probabilities were visualized in a histogram to compare the model's predictions with the observed drop behavior.

```{r}
# --- Filter data for M. mediator (mm) and Drop vs No Drop ---
mediator_data_drop <- d0 %>% 
  filter(wasp.species == "mm") %>%
  select(host.weight, thrash.frequency, host.drop) %>%
  mutate(Drop = ifelse(host.drop == 1, "Drop", "No Drop"))

# --- Logistic Regression Model: Drop vs No Drop ---
# Logistic regression model for Drop vs No Drop
logit_model_drop <- glm(host.drop ~ host.weight + thrash.frequency, 
                        data = mediator_data_drop, family = binomial(link = "logit"))

# Displaying the summary of the logistic regression model
summary(logit_model_drop)

# Predicted probabilities from the Drop vs No Drop model
mediator_data_drop$predicted_prob <- predict(logit_model_drop, type = "response")

# Plotting predicted probabilities vs actual Drop status
ggplot(mediator_data_drop, aes(x = predicted_prob, fill = Drop)) +
  geom_histogram(position = "dodge", binwidth = 0.05, alpha = 0.7) + 
  labs(title = "Predicted Probability of Drop vs Actual Drop for M. mediator",
       x = "Predicted Probability", y = "Frequency") +
  scale_fill_manual(values = c("Drop" = "blue", "No Drop" = "red")) +
  theme_minimal()
```

# A logistic regression analysis was conducted to evaluate the relationship between host weight and thrashing frequency as predictors of dropping behavior (Drop vs No Drop) in M. purchicronis. The model estimated the likelihood of dropping, and predicted probabilities were visualized in a histogram to compare the model's predictions with the observed drop behavior.

```{r}
# --- M. pulchricornis (mp) - Drop vs No Drop ---
# Filter data for M. pulchricornis (mp) and select relevant columns
pulchricornis_data_drop <- d0 %>% 
  filter(wasp.species == "mp") %>%
  select(host.weight, thrash.frequency, host.drop)

# Logistic regression model for Drop vs No Drop
logit_model_drop_mp <- glm(host.drop ~ host.weight + thrash.frequency, 
                           data = pulchricornis_data_drop, 
                           family = binomial(link = "logit"))

# Display summary of the logistic regression model
summary(logit_model_drop_mp)

# Predicted probabilities from the logistic regression model
pulchricornis_data_drop$predicted_prob <- predict(logit_model_drop_mp, type = "response")

# Plot predicted probabilities vs actual Drop status
ggplot(pulchricornis_data_drop, aes(x = predicted_prob, fill = as.factor(host.drop))) +
  geom_histogram(position = "dodge", binwidth = 0.05, alpha = 0.7, color = "black") +
  labs(title = "Predicted Probability of Drop for M. pulchricornis",
       x = "Predicted Probability", y = "Frequency", fill = "Drop Status") +
  scale_fill_manual(values = c("0" = "blue", "1" = "red"), 
                    labels = c("No Drop", "Drop")) +
  theme_minimal()
```
