---
title: "Defensive Behaviour of M.separata in Response to Different parasitoids species"
author: "BlessinG_Amira_Aizada_Sami_ALi"
format: html
editor: visual
---

# Loading the packages

```{r}
# Load necessary libraries
library(ggplot2)   # For plotting
library(dplyr)     # For data manipulation
library(readxl)    # For reading Excel files
library(sandwich)  # For robust standard errors
library(lmtest)    # For robust standard errors
```

# Extracting the data

```{r}
# Set file path (replace with your actual file path)
file_path <- "data/parasitoid.xlsx"

d0 <- read_excel(file_path)
```

# **This script analyzes the relationship between host weight, thrashing frequency, and escape probability for *M. mediator* using logistic regression.**

```{r}
# --- Filtering d0 for M. mediator (mm) and Escape vs No Escape ---
mediator_d0_escape <- d0 %>% 
  filter(wasp.species == "mm") %>%
  select(host.weight, thrash.frequency, host.escape) %>%
  mutate(Escape = ifelse(host.escape == 1, "Escape", "No Escape"))

# --- Logistic Regression Model: Escape vs No Escape ---
# GLM for logistic regression (binary outcome: Escape vs No Escape)
logit_model_escape <- glm(host.escape ~ host.weight + thrash.frequency, 
                          data = mediator_d0_escape, family = binomial(link = "logit"))

# --- Compute Robust Standard Errors ---
robust_se <- sqrt(diag(vcovHC(logit_model_escape, type = "HC3")))  # Using HC3 for robust standard errors

# Display the results with robust standard errors
coeftest(logit_model_escape, vcov = vcovHC(logit_model_escape, type = "HC3"))

# --- Predicted probabilities from the logistic regression model ---
mediator_d0_escape$predicted_prob <- predict(logit_model_escape, type = "response")

# --- Plot predicted probabilities vs actual Escape status ---
ggplot(mediator_d0_escape, aes(x = predicted_prob, fill = Escape)) +
  geom_histogram(position = "dodge", binwidth = 0.05, alpha = 0.7) + 
  labs(title = "Predicted Probability of Escape vs Actual Escape for M. mediator",
       x = "Predicted Probability", y = "Frequency") +
  scale_fill_manual(values = c("Escape" = "blue", "No Escape" = "red")) +
  theme_minimal()
```

# A logistic regression analysis was conducted to assess the relationship between host weight and thrashing frequency as predictors of escape behavior (Escape vs No Escape) for M. pulchricornis.

```{r}
#M. pulchricornis (mp) - Escape vs No Escape ---
# Filter data for M. pulchricornis (mp) and select relevant columns
pulchricornis_data_escape <- d0 %>% 
  filter(wasp.species == "mp") %>%
  select(host.weight, thrash.frequency, host.escape)

# Logistic regression model for Escape vs No Escape
logit_model_escape_mp <- glm(host.escape ~ host.weight + thrash.frequency, 
                             data = pulchricornis_data_escape, 
                             family = binomial(link = "logit"))

# Display summary of the logistic regression model
summary(logit_model_escape_mp)

# Predicted probabilities from the logistic regression model
pulchricornis_data_escape$predicted_prob <- predict(logit_model_escape_mp, type = "response")

# Plot predicted probabilities vs actual Escape status
ggplot(pulchricornis_data_escape, aes(x = predicted_prob, fill = as.factor(host.escape))) +
  geom_histogram(position = "dodge", binwidth = 0.05, alpha = 0.7, color = "black") +
  labs(title = "Predicted Probability of Escape for M. pulchricornis",
       x = "Predicted Probability", y = "Frequency", fill = "Escape Status") +
  scale_fill_manual(values = c("0" = "red", "1" = "blue"), 
                    labels = c("No Escape", "Escape")) +
  theme_minimal()
```
